---
title: "lionfish_GLM"
author: "GG"
date: "12/17/2018"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library(ggplot2)
library(tidyverse)
library(readxl)
library(janitor)
library(dbplyr)
```

```{r}
darwin <- read_csv("~/Documents/darwin_stats2.csv")
```

```{r}
library(vegan)
```

Visualizing the GLM resutls with a PCA analysis
```{r}
# remove unwanted variables 
darwin.pca <- darwin[, 2:8]
darwin.pca[6] = NULL
darwin.pca[5] = NULL
darwin.den <- (darwin[, 1] +1)
 
# apply PCA - scale. = TRUE is highly 
# advisable, but default is FALSE. 
pca <- prcomp(darwin.pca,
                 center = TRUE,
                 scale. = TRUE) 
```

```{r}
# print method
print(pca)
```

In order to create the biplot we need devtools and ggbiplot available from CRAN and github

```{r}
library(devtools)
library(ggbiplot)
library(ggfortify)
```


```{r}
d.factanal2 <- factanal(darwin.pca, factors = 2, scores = 'regression')
pca.plot2 <- autoplot(d.factanal2, data = darwin, colour = 'lionfish_den', size = 4,
         loadings = TRUE, loadings.label = TRUE, loadings.label.size  = 5) + theme_classic()
pca.plot2
```
```{r}
ggsave(darwin, plot = pca.plot2, device = "eps", path = "~/Desktop/Lionfish")
```

NOW FOR THE GLM

GLM model with all variables included showing covariance among possible drivers. 
```{r}
scale((darwin[,2:8]), center = TRUE, scale = TRUE)
fit3 <- glm((lionfish_den + 1)~fish_bio+fish_den+p_furc+fish_div+temp, data=darwin, family = Gamma())
summary(fit3)
```
Testing covairance
```{r}
library(car)
sqrt(vif(fit3))
```
*using Gamma distribution as this model results in the most "normal" data, see plots below of model fits*  
```{r}
scale((darwin[,2:8]), center = TRUE, scale = TRUE)
fit_bio <- glm((lionfish_den + 1)~fish_bio, data=darwin, family = Gamma)
fit_den <- glm((lionfish_den + 1)~fish_den, data=darwin, family = Gamma)
fit_div <- glm((lionfish_den + 1)~fish_div, data=darwin,family = Gamma, start = c(0.5, 0.5))
fit_pfur <- glm((lionfish_den + 1)~p_furc, data=darwin,family = Gamma)
fit_temp <- glm((lionfish_den + 1)~temp, data=darwin, family = Gamma)
AIC(fit_bio, fit_den, fit_pfur, fit_temp, fit_div)
coef(fit_bio, fit_den, fit_pfur, fit_temp, fit_div)
summary(fit_bio)
summary(fit_den)
summary(fit_div)
summary(fit_pfur)
summary(fit_temp)
```

```{r}
hist(resid(fit_temp))
qqnorm(resid(fit_temp))
qqline(resid(fit_temp))
```

```{r}
hist(resid(fit_den))
qqnorm(resid(fit_den))
qqline(resid(fit_den))
```

```{r}
hist(resid(fit_div))
qqnorm(resid(fit_div))
qqline(resid(fit_div))
```

```{r}
hist(resid(fit_bio))
qqnorm(resid(fit_bio))
qqline(resid(fit_bio))
```

```{r}
hist(resid(fit_pfur))
qqnorm(resid(fit_pfur))
qqline(resid(fit_pfur))
```

```{r}
coef(fit_bio)
coef(fit_pfur) 
coef(fit_temp) 
coef(fit_div)
coef(fit_den)
```
